Tree size, microhabitat, and hydraulic traits (interactively) shape drought response in a temperate broadleaf forest

### 2019-08-06 manuscript/climate_sensitivity_draft.Rmd
Temporary update

Here, we combine tree-ring records covering three droughts (1964-66, 1977, 1999), species functional and hydraulic trait measurments, and forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to test a series of hypotheses and associated specific predictions (Table 1) designed to yield functional understanding of how tree size, microenrivonment, and species' traits collectively shape drought responses. First, we focus on the role of tree size and its interaction with microenvironement. We confirm that, consistent with most forests globally [@bennett_larger_2015], larger-diameter trees have lower drought resistance in this forest, which is in an ecoregion represented by only one study in Bennett et al (2015) (H1.0). We then test hypotheses designed to disentangle the relative importance of tree height (H1.1), crown exposure (H1.2), and root water access, which should be greater for larger trees indry but not perpetually wet microsites (H1.3). Second, we focus on the role of species' functional and hydraulic traits and their interaction with tree height. We hypothesize that drought resistance will follow predicted and observed patterns in relation to wood density, specific leaf area, and ring porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area shrinkage upon dehydration and turgor loss point will prove better predictors (P2.1d-e). We then test whether these traits correlate with tree height (P2.2), potentially driving the observed tendency for taller trees to suffer more during drought (P2.3). Finally, we focused on variability among droughts, asking how  community resistance varied across droughts (H3.1) and whether the factors confirming vulnerability or resistance varied across droughts (H3.2).

**Table 1. Summary of hypotheses, predictions and results** We count predictions as fully supported (or rejected) when the direction of response matches (or contradicts) the prediction and model dAIC>=2 relative to the approriate null model or to any alternative multivariate model with within 2 dAIC. Parentheses indicate that predictions were partially supported (or rejected)-- i.e., with direction of response matching (or contradicting) the prediction but model dAIC<2 relative to the approriate null model, when response partially matched predicted order for categorical variable (canopy position), or when an alternative multivariate model within 2 dAIC had a contradictory response. "(yes/no) or ("no/yes")" indicates tendancies in opposite directions in univariate tests versus the best full models, respectively. (Table 1.xlsx)

![Table 1](tables_figures/Table1.png){width=500px}

@@

### Results
Once the data was collected, linear mixed models were run following the order of the hypotheses as seen in Figure ??? [individual_tested_traits]. Using the [@R-pointRes] package, we set up models with the resistance value as the response variable, and each prediction's variable as the independent variable. Variables' importance in predicting drought tolerance was calculated from mixed-effects models and the lowest AICc [@R-lme4, @R-AICcmodavg].Null models were determined in order of the predictions. First, we analyzed the combined scenario to determine if "year" was significant. Upon establishing this, we tested height and DBH as size parameters. Although both were significant, height was kept due to its larger delta AICc compared with the null model. We then tested the remaining biophysical and hydraulic traits individually against a null model containing height and year. This yielded Figure ??? (cand_full). All variables with dAICc >2 were used as candidates for each scenario's best model (figure ???? (tested_traits_best))

>>>>

Here, we combine tree-ring records covering three droughts (1964-66, 1977, 1999), species functional and hydraulic trait measurments, and forest census data from a 25.6-ha ForestGEO plot in Virginia (USA) to test a series of hypotheses and associated specific predictions (Table 1) designed to yield functional understanding of how tree size, microenrivonment, and species' traits collectively shape drought responses. 
First, we focus on the role of tree size and its interaction with microenvironement. 
We confirm that, consistent with most forests globally [@bennett_larger_2015], larger-diameter trees have lower drought resistance in this forest, which is in an ecoregion represented by only one study in Bennett et al (2015) (H1.0). 
We then test hypotheses designed to disentangle the relative importance of tree height (H1.1), crown exposure (H1.2), and root water access, which should be greater for larger trees in dry but not perpetually wet microsites (H1.3). 
Second, we focus on the role of species' functional and hydraulic traits and their interaction with tree height. We hypothesize that drought resistance will follow predicted and observed patterns in relation to wood density, specific leaf area, and ring porosity (P2.1a-c), but that leaf hydraulic traits such as leaf area shrinkage upon dehydration and turgor loss point will prove better predictors (P2.1d-e). 
We then test whether these traits correlate with tree height (P2.2), potentially driving the observed tendency for taller trees to suffer more during drought (P2.3). 
Finally, we focused on variability among droughts, asking how  community resistance varied across droughts (H3.1) and whether the factors confirming vulnerability or resistance varied across droughts (H3.2).

@@

### Results
![Table 1](tables_figures/Table1.png){width=500px}

**Table 1. Summary of hypotheses, predictions and results** 
We count predictions as fully supported (or rejected) when the direction of response matches (or contradicts) the prediction. 
H2.3-H3.2 are based on if models containing species traits in H2.1 had dAIC>=2 relative to the appropriate null model or to any alternative multivariate model within 2 dAIC. 
Parentheses indicate that predictions were partially supported (or rejected). In other words, parentheses indicate when the direction of response matched (or contradicted) the prediction in some but not all models with dAIC<2 relative to the appropriate null model. With categorical variables such as crown position,  a "(yes)" notation describes when the trend matched the prediction, but it wasn't significant 
A "(yes/no) or ("no/yes")" indicates tendencies in opposite directions in univariate tests versus the best full models, respectively. 
"(yes)
(Table 1.xlsx)

trying to test how the negative effect of height is caused by the traits,

(no) means there's a trend in the expected direction but not consistently significant

H2.3
if the effect is negative and is in every top model (<dAIC2), that would be "yes."
if the effect is negative mostly

Once the data was collected, linear mixed models were run following the order of the hypotheses as seen in Figure ??? [individual_tested_traits]. Using the [@R-pointRes] package, we set up models with the resistance value as the response variable, and each prediction's variable as the independent variable. Variables' importance in predicting drought tolerance was calculated from mixed-effects models and the lowest AICc [@R-lme4, @R-AICcmodavg].Null models were determined in order of the predictions. First, we analyzed the combined scenario to determine if "year" was significant. Upon establishing this, we tested height and DBH as size parameters. Although both were significant, height was kept due to its larger delta AICc compared with the null model. We then tested the remaining biophysical and hydraulic traits individually against a null model containing height and year. This yielded Figure ??? (cand_full). All variables with dAICc >2 were used as candidates for each scenario's best model (figure ???? (tested_traits_best))
